### 2021-02-18 text/draft.tex
reconciliation

\begin{figure}
\centering
\includegraphics{images/keywords_clean_histogram_abc.pdf}
\caption{In \texttt{keywords/keywords\_clean\_histogram.json}, the
distribution of keywords is strongly skewed right
(\(min = Q1 = M = Q3 = 1\) and \(max = 910\)). However, even though
keywords with only one occurrence constitute over 75\% of the total
keywords, they constitute less than 35\% of the total occurrences. The
most common keywords with 50 or more occurrences are extreme outliers
but make up almost 20\% of the total occurrences.}
\end{figure}

@@

As explained in section X, Annif assigns index terms from a controlled
vocabulary. If we want to assess the quality of the indexing via a gold
standard, we must therefore ensure that the gold standard makes use of
the vocabulary used by Annif. As seen in section Y, the relevant
(English) vocabularies are Wikidata YSO. The next step in constructing
the native gold standard is hence to reconcile the extracted and cleaned
keywords with YSO and Wikidata.

The tool of choice for reconciliation is OpenRefine (see
\protect\hyperlink{openrefine}{section OpenRefine}). We import
\texttt{keywords/keywords\_clean\_histogram.json} as a new project.

>>>>

\begin{figure}
\centering
\includegraphics{images/keywords_clean_histogram_abc.pdf}
\caption{In \texttt{keywords/keywords\_clean\_histogram.json}, the
distribution of keywords is strongly skewed right
(\(min = Q1 = M = Q3 = 1\) and \(max = 910\)). However, even though
keywords with only one occurrence constitute over 75\% of the total
keywords, their occurrences constitute less than 35\% of the total
occurrences. The most common keywords with 50 or more occurrences are
extreme outliers but make up almost 20\% of the total occurrences.}
\end{figure}

@@

As explained in section X, Annif assigns index terms from a controlled
vocabulary. If we want to assess the quality of the indexing via a gold
standard, we must therefore ensure that the gold standard makes use of
the vocabulary used by Annif. As seen in section Y, the relevant
(English) vocabularies are Wikidata YSO. The next step in constructing
the native gold standard is hence to match the extracted and cleaned
keywords with keywords from YSO and Wikidata. This process is called
``reconciliation'' (see https://docs.openrefine.org/manual/reconciling)
and the tool of choice for this task is OpenRefine (see
\protect\hyperlink{openrefine}{section OpenRefine}).

In what follows, I will describe how the cleaned keywords were
reconciled with Wikidata and YSO, and which additional steps for
refinement were undertaken. In total, 2'104 data transformation
operations were performed; the complete operation history is available
as \texttt{/keywords/operation\_history.json}. First
\texttt{keywords/keywords\_clean\_histogram.json} was set as a new
project in OpenRefine. The \texttt{keyword} column was then duplicated
and reconciled with Wikidata. Here the parameters were chosen as
follows: reconcile against no particular type, and auto-match candidates
with high confidence.
